---
title: "JHU Statistical Inference - Simulations"
author: "LiJun"
date: "24 October 2018"
output:
  word_document: default
  pdf_document: default
---

###Introduction
#### This analysis aims to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

### Simulations
```{r}
#Set simulation variables lambda, exponentials an seed
set.seed(1234)
lambda = 0.2
exponentials = 40

#Run Simulations with variables
simulations = NULL
for (i in 1 : 1000) simulations = c(simulations, mean(rexp(exponentials, lambda)))
```

###1.Show the sample mean and compare it to the theoretical mean of the distribution.
```{r}
#Sample mean - mean of the simulations
mean(simulations)

#Theoretical mean - lambda^-1
lambda^-1

#Comparison between sample mean and theoretical mean
compare_mean = abs(mean(simulations)-lambda^-1)
compare_mean
```
####Conclusion: There is only a slight difference between the sample mean and the inference mean.

###2.Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r}
#Sample variance - variance of the simulations
var(simulations)

#Theoretical mean - (lambda * sqrt(n))^-2
(lambda * sqrt(exponentials))^-2

#Comparison between sample variance and theoretical variance
compare_variance = abs(var(simulations)-(lambda * sqrt(exponentials))^-2)
compare_variance
```
####Conclusion: There is only a slight difference between the sample variance and the inference variance.

###3.Show that the distribution is approximately normal.
```{r}
#Distribution - based on density histogram of 1000 simulations, overlay with theoretical normal distribution 
library(ggplot2)
ggplot(data.frame(y=simulations), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="blue", color="black") +
  stat_function(fun=dnorm, args=list(mean=lambda^-1, 
                                    sd=(lambda*sqrt(exponentials))^-1), 
                size=2) +
  labs(title="Plot of the Simulations", x="Simulation Mean")
```